---
title: "Project 1 Analysis"
author: "Group 6"
date: "4/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

source("aim2.R")
#########################################################
## Libraries
#########################################################
library(ctmle)
library(DynTxRegime)
library(glmnet)
library(tidyverse)
library(magrittr)
library(readxl)
library(lubridate)
library(here)
library(table1)
library(sandwich)
library(metaviz)
library(randomForest)
library(earth)
here <- here::here

#########################################################
## Data
#########################################################

# You will need to change this to your path
# So that we don't have to comment out other people's directories, I'm putting some conditional logic
cur.directory <- here()
if(str_detect(string = cur.directory, pattern = "jsperger")){
  data.dir.path <- "/Users/jsperger/Dropbox/Consulting Project/Data/"
}
if(str_detect(string = cur.directory, pattern = "nikki")){
  data.dir.path <- "Fix me/Dropbox/Consulting Project/Data/"
}
if(str_detect(string = cur.directory, pattern = "haley")){
  data.dir.path <- "Fix me/Dropbox/Consulting Project/Data/"
}
if(str_detect(string = cur.directory, pattern = "michael")){
  data.dir.path <- "Fix me/Dropbox/Consulting Project/Data/"
}
if(exists("data.dir.path") == FALSE | str_detect(data.dir.path, "Fix me") == TRUE) stop("Fix data directory specification")

clti <- read_csv(file = paste0(data.dir.path, "analyticData.csv"),
                 col_types = cols(
                   ptNum = col_character(), 
                   ptSide = col_character(), 
                   GENDER = col_factor(), 
                   PATIENT_RACE_1 = col_factor(), 
                   ETHNICITY = col_factor(), 
                   ageAtIndexProcedure = col_integer(), 
                   primaryProcedure = col_factor(), 
                   CLI_DX_DATE = col_datetime(), 
                   hemodynamicLaterality = col_factor(), 
                   InflowDisease = col_integer(), 
                   OutflowDisease = col_integer(), 
                   RunoffDisease = col_integer(), 
                   stentBPG = col_integer(), 
                   maxRutherfordClass = col_factor(ordered = TRUE), 
                   ischemia = col_factor(ordered = TRUE), 
                   woundClass = col_factor(ordered=TRUE), 
                   CHRONIC_PULM_DISEASE = col_integer(), 
                   COAGULOPATHY = col_integer(), 
                   `CONGESTIVE HEART FAILURE-COMPLICATED` = col_integer(), 
                   CHF = col_integer(), 
                   `DEMENTIA-NA` = col_integer(), 
                   `DIABETES-COMPLICATED` = col_integer(), 
                   DIABETES = col_integer(), 
                   `DIABETES-UNCOMPLICATED` = col_integer(), 
                   `HYPERTENSION-COMPLICATED` = col_integer(), 
                   `HYPERTENSION-UNCOMPLICATED` = col_integer(), 
                   MI = col_integer(), 
                   OBESITY = col_integer(), 
                   `RENAL DISEASE-CKD` = col_integer(), 
                   `RENAL DISEASE-COMPLICATED` = col_integer(), 
                   `RENAL DISEASE-ESRD` = col_integer(), 
                   RENAL_DISEASE = col_integer(), 
                   SMOKING = col_integer(), 
                   VENOUS_INSUFFICIENCY = col_integer(), 
                   WEIGHT_LOSS = col_integer(), 
                   CEREBROVASCULAR = col_integer(), 
                   VTE = col_integer(), 
                   CAD = col_integer(), 
                   anyAnemia = col_integer(), 
                   anyHyperlipidemia = col_integer(), 
                   mort2yr = col_integer(), 
                   majorAmp2yr = col_integer(), 
                   mace2yr = col_integer(), 
                   ampFreeSurv2yr = col_integer() 
                 ))

# This is for an easier to read table1
# temporary, can remove later
clti.binary.as.logical <- read_csv(file = paste0(data.dir.path, "analyticData.csv"),
                 col_types = cols(
                   ptNum = col_character(), 
                   ptSide = col_character(), 
                   GENDER = col_factor(), 
                   PATIENT_RACE_1 = col_factor(), 
                   ETHNICITY = col_factor(), 
                   ageAtIndexProcedure = col_integer(), 
                   primaryProcedure = col_factor(), 
                   CLI_DX_DATE = col_datetime(), 
                   hemodynamicLaterality = col_factor(), 
                   InflowDisease = col_logical(), 
                   OutflowDisease = col_logical(), 
                   RunoffDisease = col_logical(), 
                   stentBPG = col_logical(), 
                   maxRutherfordClass = col_factor(ordered = TRUE), 
                   ischemia = col_factor(ordered = TRUE), 
                   woundClass = col_factor(ordered=TRUE), 
                   CHRONIC_PULM_DISEASE = col_logical(), 
                   COAGULOPATHY = col_logical(), 
                   `CONGESTIVE HEART FAILURE-COMPLICATED` = col_logical(), 
                   CHF = col_logical(), 
                   `DEMENTIA-NA` = col_logical(), 
                   `DIABETES-COMPLICATED` = col_logical(), 
                   DIABETES = col_logical(), 
                   `DIABETES-UNCOMPLICATED` = col_logical(), 
                   `HYPERTENSION-COMPLICATED` = col_logical(), 
                   `HYPERTENSION-UNCOMPLICATED` = col_logical(), 
                   MI = col_logical(), 
                   OBESITY = col_logical(), 
                   `RENAL DISEASE-CKD` = col_logical(), 
                   `RENAL DISEASE-COMPLICATED` = col_logical(), 
                   `RENAL DISEASE-ESRD` = col_logical(), 
                   RENAL_DISEASE = col_logical(), 
                   SMOKING = col_logical(), 
                   VENOUS_INSUFFICIENCY = col_logical(), 
                   WEIGHT_LOSS = col_logical(), 
                   CEREBROVASCULAR = col_logical(), 
                   VTE = col_logical(), 
                   CAD = col_logical(), 
                   anyAnemia = col_logical(), 
                   anyHyperlipidemia = col_logical(), 
                   mort2yr = col_logical(), 
                   majorAmp2yr = col_logical(), 
                   mace2yr = col_logical(), 
                   ampFreeSurv2yr = col_logical() 
                 ))


clti.base <- read.csv(file = paste0(data.dir.path, "analyticData.csv"))
# Use base R naming scheme for column names. Saves some headaches later with DynTxRegime
names(clti) <- names(clti.base)
#clti <- fastDummies::dummy_cols(clti %>% select(-all_of(id.var.names)))

rm(clti.base)
```

```{r prevalence_summary}
clti.binary.as.logical %>% select(-all_of(c("ptNum", "ptSide", "CLI_DX_DATE"))) %>% table1(x = ~. | primaryProcedure, data = .)
```

```{r vars}
id.var.names <- c("ptNum", "ptSide", "CLI_DX_DATE")

outcome.var.names <- c("ampFreeSurv2yr",
                       "majorAmp2yr",
                       "mort2yr",
                       "mace2yr")

disease.var.names <- c("maxRutherfordClass",
                  "woundClass",
                  "ischemia",
                  "InflowDisease",
                  "OutflowDisease",
                  "RunoffDisease",
                  "hemodynamicLaterality", #Not 100% this belongs here JS
                  "stentBPG" # Not sure about this one either
                  )

comorbidity.var.names <- c("anyAnemia",
                           "CEREBROVASCULAR",
                           "CHRONIC_PULM_DISEASE",
                           "COAGULOPATHY",
                           "CONGESTIVE HEART FAILURE-COMPLICATED",
                           "CHF",
                           "DEMENTIA-NA",
                           "DIABETES-COMPLICATED",
                           "DIABETES",
                           "DIABETES-UNCOMPLICATED",
                           "anyHyperlipidemia",
                           "HYPERTENSION-COMPLICATED",            
                           "HYPERTENSION-UNCOMPLICATED",
                           "MI",
                           "OBESITY",
                           "RENAL DISEASE-CKD",
                           "RENAL DISEASE-COMPLICATED",
                           "RENAL DISEASE-ESRD",
                           "RENAL_DISEASE",
                           "SMOKING",
                           "VENOUS_INSUFFICIENCY",
                           "WEIGHT_LOSS",
                           "VTE",
                           "CAD") %>% 
  str_replace_all(., pattern = "[:space:]|-", replacement= ".")

demo.var.names <- c("GENDER",
                    "PATIENT_RACE_1",
                    "ETHNICITY",
                    "ageAtIndexProcedure")


intervention.var.name <- "primaryProcedure"

# This stuff was just for trying to make sure I had the variable names right 
#ided.vars <- c(id.var.names, outcome.var.names, comorbidity.var.names, disease.var.names, 
#               demo.var.names, intervention.var.name)
#unknown.vars <- names(clti)[!names(clti) %in% ided.vars]
```

```{r preprocessing}
#########################################################
## Preprocessing
#########################################################
# Create a nonwhite indicate
# recode a few factors as 0-1 integers
# create a ampFree where 1 is bad and 0 is good to be in line with the other outcome vars
clti %<>%  mutate(primaryTreatmentInt = if_else(primaryProcedure == "WM", 1, 0),
                  nonWhite = if_else(PATIENT_RACE_1 == "WHITE OR CAUCASIAN", 1, 0),
                  ampMort2yr = abs(ampFreeSurv2yr-1)) %>% 
          mutate_at(.vars = "hemodynamicLaterality", ~if_else(. == "Right" | . == 1, 1, 0)) %>% 
  mutate_at(.vars = "GENDER", ~if_else(. == "FEMALE"| . == 1, 1, 0))

prevalence.prop.tables <- clti %>% select(-any_of(c("ptNum", "ptSide", "CLI_DX_DATE", "ageAtIndexProcedure", outcome.var.names, "ampMort2yr"))) %>% 
  apply(., 2, function(x){
    temp.data <- data.frame(clti$primaryProcedure, x)
    round(prop.table(table(temp.data), 1),2)
  })
#TODO: 
# This is really checking whether the prevalence is between .18-.82
prev.abv.cutoff <- map_lgl(prevalence.prop.tables, ~all(. >= .18))
nonnum.covars.to.include <- names(prev.abv.cutoff)[prev.abv.cutoff]

```

The non-numeric variables to be considered based on having prevalence greater than 20% in both intervention groups are

```{r prev_vars}
print(nonnum.covars.to.include)

```

# Aim 2

```{r aim2_model_fitting, cache=TRUE, warning=FALSE, message=FALSE}
covar.names <- c("ageAtIndexProcedure", nonnum.covars.to.include)

set.seed(8675309)
# Tried 10-fold and got a funky error
ampsurv.ctmle.fit <- fitCTMLE(clti, "ampMort2yr", covar.names,
                              SL.library = c("SL.glm", "SL.glmnet", "SL.randomForest", 
                                            "SL.ksvm", "SL.earth"), 
                              gbound = .05)
mort.ctmle.fit <- suppressWarnings(fitCTMLE(clti, "mort2yr", covar.names,
                           SL.library = c("SL.glm", "SL.glmnet", "SL.randomForest", 
                                          "SL.ksvm", "SL.earth"), 
                           gbound = .05))
amp.ctmle.fit <- suppressWarnings(fitCTMLE(clti, "majorAmp2yr", covar.names, 
                          SL.library = c("SL.glm", "SL.glmnet", "SL.randomForest",
                                         "SL.ksvm", "SL.earth"), 
                          gbound = .05))
mace.ctmle.fit <- suppressWarnings(fitCTMLE(clti, "mace2yr", covar.names, 
                           SL.library = c("SL.glm", "SL.glmnet", "SL.randomForest", 
                                          "SL.ksvm", "SL.earth"), 
                           gbound = .05))

ampsurv.aipw.fit <- fitAIPWE(clti, "ampMort2yr", covar.names)
mort.aipw.fit <- fitAIPWE(clti, "mort2yr", covar.names)
amp.aipw.fit <- fitAIPWE(clti, "majorAmp2yr", covar.names)
mace.aipw.fit <- fitAIPWE(clti, "mace2yr", covar.names)
```

```{r aim2_plot}
ctmle.fits <- list("Amp_mort_2yr-CTMLE" = ampsurv.ctmle.fit, 
                  "Mort_2yr-CTMLE" = mort.ctmle.fit, 
                  "Amp_2yr-CTMLE" = amp.ctmle.fit, 
                  "MACE_2yr-CTMLE" = mace.ctmle.fit)
ctmle.ests <- map_dfr(ctmle.fits, ~summary(.)[c("est", "var")] %>% unlist()) %>% t()
ctmle.ests[,2] <- sqrt(ctmle.ests[,2])
# Use a linear model and sandich estimator to estimate the risk difference
ampsurv.unadj <- lm(ampMort2yr ~ 1 + primaryTreatmentInt, data = clti)
ampsurv.sandwich.se <- diag(sandwich(ampsurv.unadj))^.5
mort.unadj <- lm(mort2yr ~ 1 + primaryTreatmentInt, data = clti)
mort.sandwich.se <- diag(sandwich(mort.unadj))^.5
amp.unadj <- lm(majorAmp2yr ~ 1 + primaryTreatmentInt, data = clti)
amp.sandwich.se <- diag(sandwich(amp.unadj))^.5
mace.unadj <- lm(mace2yr ~ 1 + primaryTreatmentInt, data = clti)
mace.sandwich.se <- diag(sandwich(mace.unadj))^.5

unadj.results <- rbind(c(coef(ampsurv.unadj)[2], ampsurv.sandwich.se[2]),
                       c(coef(mort.unadj)[2], mort.sandwich.se[2]),
                       c(coef(amp.unadj)[2], amp.sandwich.se[2]),
                       c(coef(mace.unadj)[2], mace.sandwich.se[2]))

results.sum.df <- as.data.frame(rbind(unadj.results, ctmle.ests))
names(results.sum.df) <- c("RD.hat", "se")
results.sum.df$Outcome <- c("Amp or Death", "Death", "Amputation", "MACE", "Ampu or Death", "Death", "Amputation", "MACE")
results.sum.df$Method <- c(rep("Unadjusted", 4), rep("CTMLE", 4))
results.sum.df %<>% arrange(Outcome)
results.sum.df %<>% mutate(Lower = RD.hat - 1.96*se,
                           Upper = RD.hat + 1.96*se)

aim2.plot <- suppressWarnings(ggplot(data = results.sum.df, aes(x = Outcome, y = RD.hat,  ymin = Lower, ymax = Upper))+
  geom_pointrange(aes(col = Method)) +
  geom_hline(aes(fill=Group),yintercept =0, linetype=2)+
    xlab('Outcome')+ ylab("Risk Difference (95% Confidence Interval)")+
    geom_errorbar(aes(ymin=Lower, ymax=Upper,col=Method),width=0.5,cex=1)+ 
    facet_wrap(~Outcome,strip.position="left",nrow=9,scales = "free_y") +
    theme(plot.title=element_text(size=16,face="bold"),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_text(size=12,face="bold"),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold"))+
    coord_flip())
```

```{r}
aim2.plot
```

#### CTMLE Covariates

```{r}
cat("Amputation-free survival Covariates: ", extractCTMLEFinalCovar(ampsurv.ctmle.fit))

cat("Amputation Covariates: ", extractCTMLEFinalCovar(amp.ctmle.fit))

cat("Mortality Covariates: ", extractCTMLEFinalCovar(mort.ctmle.fit))

cat("MACE Covariates: ", extractCTMLEFinalCovar(mace.ctmle.fit))
```
#### CTMLE Estimates

```{r}
print("Amputation or mortality")
ampsurv.ctmle.fit

print("Amputation")
amp.ctmle.fit

print("Mortality")
mort.ctmle.fit

print("Mace")

mace.ctmle.fit
```

## Aim 2 Model Dump

Notice: Amputation-free survival has been recoded so 1 is bad and 0 is good for the moment so that it is in line with the other outcome variables where 1 is bad and 0 is good. Easier to keep track of for the moment while trying to diagnose what's happening. 

```{r aim2_summaries}
aim2.mods <- list("Amp_mort_2yr-CTMLE" = ampsurv.ctmle.fit, 
                  "Mort_2yr-CTMLE" = mort.ctmle.fit, 
                  "Amp_2yr-CTMLE" = amp.ctmle.fit, 
                  "MACE_2yr-CTMLE" = mace.ctmle.fit,
                  "Amp_mort_2yr-AIPW" = ampsurv.aipw.fit, 
                  "Mort_2yr-AIPW" = mort.aipw.fit, 
                  "Amp_2yr-AIPW" = amp.aipw.fit, 
                  "MACE_2yr-AIPW" = mace.aipw.fit)

map(aim2.mods, ~summary(.))
```