---
title: 'Consulting project 1: Data preparation'
output:
  html_document:
    df_print: paged
---

Load packages
```{r, message = FALSE}
library(tidyverse)
library(lubridate)
```

Load the data
```{r, message = FALSE}
grid <- readxl::read_xlsx("/Users/Nikki/Dropbox/Consulting project/Data/grid2NoMRN.xlsx")
mort <- readxl::read_xlsx("/Users/Nikki/Dropbox/Consulting project/Data/mortNoMRN.xlsx")
proc <- readxl::read_xlsx("/Users/Nikki/Dropbox/Consulting project/Data/procGroupedNoMRN.xlsx")
procLat <- readxl::read_xlsx("/Users/Nikki/Dropbox/Consulting project/Data/procLatNoMRN.xlsx")
```



## Identify the primary WM patients and the primary revascularization patients

**Inclusion/Exclusion**: We only want to **include** the patients that are in `grid2NoMRN.xlsx` because those are the patients that had each dimension of desired information (qualifying TP/ABI and Rutherford score, imagining and extractble disease anatomy information). We are **excluding** patients that do not get any CLTI-related procedures after their CLTI diagnosis (e.g. patients that went to Duke for treatment).

  1. For the procedure file, filter out patients that aren't in `grid2NoMRN.xlsx`. 
  2. For the procedure file, filter out the procedures labeled as "other" (these are procedures unrelated to CLTI)
  3. For the procedure file, select and set the type of the variables that we need and remove the ones we don't
    * Need: `CLI_DX_DATE`, `PROCEDURE_DATE`, `DAYS_FROM_CLI_DX`, `procGroup`, `ptNum`
  4. For the procedure file, filter out procedures that occur more than 10 days before the CLTI diagnosis date

__Definition__ A patient is a primary wound management patient if they only receive wound management during the first 30 days after the index procedure after the CLI_DX_DATE. A patient is a primary revascularization patient if they receive a revascularization within the first 10 days after the index procedure after the CLI_DX_DATE.

```{r, message = FALSE}
# Get the ptNums for the unilats in grid2.xlsx
grid <- grid %>% mutate(ptSide = paste0(ptNum, hemodynamicLaterality))


# Find the primary WM and primary revasc patients
primaryProc <- procLat %>% mutate(ptSide = paste0(ptNum, laterality)) %>% 
  filter(ptSide %in% grid$ptSide) %>%  # keep patients from grid2 
  filter(procGroup != "other") %>% # procedures labeled as "other" do not apply to us
  select(ptNum, ptSide, CLI_DX_DATE, PROCEDURE_DATE, DAYS_FROM_CLI_DX, procGroup) %>% # Select the columns that matter to us
  mutate(ptNum = as.character(ptNum), # Fix the variable types (yay defense!)
         ptSide = as.character(ptSide), 
         CLI_DX_DATE = ymd(CLI_DX_DATE),
         PROCEDURE_DATE = ymd(PROCEDURE_DATE),
         DAYS_FROM_CLI_DX = as.numeric(DAYS_FROM_CLI_DX),
         procGroup = as.character(procGroup)) %>%
  filter(DAYS_FROM_CLI_DX >= 0) %>% # Remove procedures from before the CLI_DX_DATE
  group_by(ptSide) %>% 
  mutate(indexProcedureFlag = if_else(DAYS_FROM_CLI_DX == min(DAYS_FROM_CLI_DX), 1, 0), # Flag the index procedures
         indexProcedureDate = if_else(indexProcedureFlag == 1, PROCEDURE_DATE, as_date(NA)), # Define the index period
         indexProcedureDatePlus30 = if_else(indexProcedureFlag == 1, indexProcedureDate + days(30), as_date(NA)),
         indexProcedureDatePlus10 = if_else(indexProcedureFlag == 1, indexProcedureDate + days(10), as_date(NA)),
         indexProcedureDatePlus90 = if_else(indexProcedureFlag == 1, indexProcedureDate + days(90), as_date(NA))) %>%
  arrange(ptSide, DAYS_FROM_CLI_DX) %>%
  fill(indexProcedureDate, .direction = "down") %>%
  fill(indexProcedureDatePlus30, .direction = "down") %>%
  fill(indexProcedureDatePlus10,.direction = "down") %>%
  fill(indexProcedureDatePlus90, .direction = "down") %>%
  mutate(within30DaysOfIndexProcedureFlag = if_else(PROCEDURE_DATE <= indexProcedureDatePlus30, 1, 0)) %>% # Flag procedures within 30 days of index procedure
  mutate(within90DaysOfIndexProcedure = if_else(PROCEDURE_DATE <= indexProcedureDatePlus90, 1, 0)) %>%
  filter(within90DaysOfIndexProcedure == 1) %>% 
  select(ptSide, ptNum,  within30DaysOfIndexProcedureFlag, within90DaysOfIndexProcedure, procGroup) %>%
  mutate(procGroupReduced = if_else((procGroup == "Endovascular" | procGroup == "Open") & within30DaysOfIndexProcedureFlag == 1, "Revasc", "Unknown")) %>%
  dplyr::select(-within30DaysOfIndexProcedureFlag) %>%
  distinct() %>%
  pivot_wider(id_cols = c(ptSide, ptNum, procGroupReduced),
              names_from = procGroup,
              values_from = within90DaysOfIndexProcedure) %>%
  replace_na(list(Endovascular = 0,
                  Open = 0,
                  Debridement = 0,
                  `Major amputation` = 0,
                  Fasciotomy = 0, 
                  `Minor amputation` = 0)) %>%
  mutate(procGroupReduced = if_else((Debridement == 1 | `Minor amputation` == 1) & procGroupReduced == "Unknown", "WM", procGroupReduced)) %>%
  filter(procGroupReduced != "Unknown") %>%
  mutate(ones = 1) %>%
  pivot_wider(id_cols = c(ptSide, ptNum), names_from = procGroupReduced, values_from = ones) %>%
  replace_na(list(Revasc = 0,
                  WM = 0)) %>%
  mutate(primaryProcedure = if_else(Revasc == 1, "Revasc", "WM")) %>%
  select(ptSide, ptNum, primaryProcedure)
  
# Counts of primary procedures
table(primaryProc$primaryProcedure)

```

## Create the mortality outcomes

```{r, message = FALSE}
mort2yr <- left_join(primaryProc, mutate(mort, 
                              ptNum = as.character(ptNum), 
                              ptSide = paste0(ptNum, hemodynamicLaterality)), 
          by = c("ptNum", "ptSide")) %>%
  select(ptSide, ptNum, deathDate, timeToDeathFromStudyStart) %>%
  mutate(mort2yr = if_else(!is.na(timeToDeathFromStudyStart) & timeToDeathFromStudyStart <= 365*2, 1, 0))

```

## Create the amputation outcome

```{r}
amp2yr <- procLat %>% mutate(ptSide = paste0(ptNum, laterality)) %>% 
  filter(ptSide %in% grid$ptSide) %>%  # keep unilat patients in grid2
  filter(procGroup != "other") %>% # procedures labeled as "other" do not apply to us
  select(ptNum, ptSide, CLI_DX_DATE, PROCEDURE_DATE, DAYS_FROM_CLI_DX, procGroup) %>% # Select the columns that matter to us
  mutate(ptNum = as.character(ptNum), # Fix the variable types (yay defense!)
         ptSide = as.character(ptSide), 
         CLI_DX_DATE = ymd(CLI_DX_DATE),
         PROCEDURE_DATE = ymd(PROCEDURE_DATE),
         DAYS_FROM_CLI_DX = as.numeric(DAYS_FROM_CLI_DX),
         procGroup = as.character(procGroup)) %>%
  filter(DAYS_FROM_CLI_DX >= -10) %>% # Remove procedures from before the CLI_DX_DATE
  group_by(ptSide) %>%
  mutate(indexProcedureFlag = if_else(DAYS_FROM_CLI_DX == min(DAYS_FROM_CLI_DX), 1, 0), # Flag the index procedures
         indexProcedureDate = if_else(indexProcedureFlag == 1, CLI_DX_DATE, as_date(NA)), # Define the index period
         indexProcedureDatePlus30 = if_else(indexProcedureFlag == 1, CLI_DX_DATE + days(30), as_date(NA)),
         indexProcedureDatePlus10 = if_else(indexProcedureFlag == 1, CLI_DX_DATE + days(10), as_date(NA))) %>%
  arrange(ptSide, DAYS_FROM_CLI_DX) %>%
  fill(indexProcedureDate, .direction = "down") %>%
  fill(indexProcedureDatePlus30, .direction = "down") %>%
  fill(indexProcedureDatePlus10,.direction = "down") %>%
  mutate(inIndexPeriodFlag = if_else(PROCEDURE_DATE <= indexProcedureDatePlus30, 1, 0)) %>%
  filter(procGroup == "Major amputation") %>%
  mutate(timeToAmputation = PROCEDURE_DATE - indexProcedureDate) %>%
  filter(timeToAmputation <= years(2)) %>%
  select(ptNum, ptSide) %>%
  mutate(majorAmp2yr = 1) %>%
  distinct()

```

## Create the MACE outcomes
```{r}
```

## Merge the outcomes, comorbidities, and procedures
```{r}
preAnalyticData <- left_join(primaryProc, mort2yr, by = c("ptNum", "ptSide")) %>%
  left_join(amp2yr, by = c("ptNum", "ptSide")) %>%
  left_join(mutate(grid, ptNum = as.character(ptNum)), by = c("ptNum", "ptSide")) %>%
  replace_na(list(majorAmp2yr = 0, 
                  RunoffDisease = 0,
                  InflowDisease = 0,
                  OutflowDisease = 0)) %>%
  mutate(RunoffDisease = if_else(RunoffDisease == -50, 0, RunoffDisease),
         InflowDisease = if_else(InflowDisease == -50, 0, InflowDisease),
         OutflowDisease = if_else(OutflowDisease == -50, 0, OutflowDisease))
```

## Reduce down to analytic dataset
```{r}
# Select only the vars we need, replace NAs as needed
preAnalyticData2 <- preAnalyticData %>% select(ptNum, ptSide, # Patient Limb identifiers
                           primaryProcedure, # primary treatment
                           CLI_DX_DATE = STUDY_DATE...3, # CLTI diagnosis date
                           hemodynamicLaterality, InflowDisease,  # Anatomy
                           OutflowDisease, RunoffDisease,
                           stentBPG, # previous stent/bpg
                           maxRutherfordClass, ischemia, # Severity
                           woundClass, infectionClass,
                           ANEMIA:CAD, # comorbidities
                           mort2yr, majorAmp2yr)  

# Combine the rare comorbidities (e.g anemia)
preAnalyticData2 %>% ungroup() %>% summarise_at(vars(InflowDisease:CAD), sum)

# Remove the infection class
# Combine anemia
# Rm cerebrovascular disease complicated
# Combine hyperlipidemia complicated
# Remove the wound measures
# Make the 2yr amp free survival composite measure
analyticData <- preAnalyticData2 %>%
  mutate(anyAnemia = if_else(ANEMIA ==1 | `DEFICIENCY ANEMIA-NA` == 1, 1, 0),
         anyHyperlipidemia = if_else(`HYPERLIPIDEMIA-COMPLICATED` == 1 | `HYPERLIPIDEMIA-NA` == 1, 1, 0)) %>%
  select(-c(ANEMIA, `DEFICIENCY ANEMIA-NA`, `HYPERLIPIDEMIA-COMPLICATED`, `HYPERLIPIDEMIA-NA`,
            infectionClass, `CEREBROVASCULAR DISEASE-COMPLICATED`,
            `WOUND-COMPLICATED`, `WOUND-EXTREMITY`, `WOUND`)) %>%
  select(ptNum:CAD, anyAnemia, anyHyperlipidemia, mort2yr, majorAmp2yr) %>%
  mutate(ampFreeSurv2yr = if_else(mort2yr == 0 & majorAmp2yr == 0 , 1, 0))



  
readr::write_csv(x = analyticData, path = "analyticData.csv")
  
  
```